---
title: "Assignment 3"
author: "Data manipulation"
date: "Due 23 February 2024"
output: 
  html_document:
    code_folding: show
    theme: journal
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

#  ES482 R labs   
#      University of Victoria, Victoria BC Canada             

#  Assignment 3: Data manipulation

library(tidyverse)
```

## Submission
Please email me a copy of your R script for this assignment by start of class on **Friday February 23rd** with your first and last name followed by assignment 3 as the file name (e.g. 'marissa_dyck_assignment3.R')

> You should always be following best coding practices (see Intro to R module 1) but especially for assingment submissions. Please make sure each problem has its own header so that I can easily navigate to your answers and that your code is well organized with spaces as described in the best coding practices section and comments as needed. 

## 1 Remove NAs
- Read in the turtles_tidy data   
- In the same code chunk remove all rows with NAs
- Assign this new data to the environment as "turtles_no_na" 

```{r echo=FALSE, message=FALSE}

# code to read in turtles data from earlier
turtles_tidy <- read_csv('data/processed/turtles_tidy.csv') %>% 
  
  # change sex to a factor
  mutate(sex = as.factor(sex),
         sex = recode(sex, 
                       fem = 'female'))

# answer to Q1
turtles_no_na <- turtles_tidy %>% 
  
  # remove rows with NA
  na.omit()

turtles_no_na
```

## 2 Make a new variable
Using the turtles_no_na data, make a new variable called "size_class" based on the "weight" variable using `case_when()` whereby

* weights less than 4 are juvenile

* weights greater than 7 are adult

* weights between 4 and 7 are subadult

(There are multiple ways to do this which is why there are multiple printouts, but they will yield the same answer)
```{r echo=FALSE}

turtles_no_na <- turtles_no_na %>% 
  mutate(size.class = case_when(
    weight < 4 ~ "juvenile",
    weight > 7 ~ "adult",
    TRUE  ~ "subadult"
  ))

turtles_tidy$size.class

# alternatively
turtles_no_na <- turtles_no_na %>% 
  mutate(size_class = case_when(
    weight < 4 ~ "juvenile",
    weight > 7 ~ "adult",
    weight >= 4 & weight <= 7 ~ 'subadult'
  ))

turtles_no_na$size_class
```


## 3 Replace values with NA
In the **turtles_tidy data** (*not the turtles_no_na data*) replace ALL variable values for tags 104 and 105 with NAs

*Hint you will need to create a vector for the tag numbers you want to replace and use `mutate()`*
```{r echo=FALSE}
# list of tags we do not trust the data for
bad.tags <- c(104, 105)

turtles_tidy %>% 
  mutate(
    sex = replace(sex,
                  tag %in% bad.tags,
                  NA),
    c_length = replace(c_length,
                       tag %in% bad.tags,
                       NA),
    h_width = replace(h_width,
                      tag %in% bad.tags,
                      NA),
    weight = replace(weight,
                     tag %in% bad.tags,
                     NA)
  )

# or... use some more tidyverse helper functions and tricks!

turtles_tidy %>% 
  mutate(across(
    c("sex","c_length","h_width", "weight"),
    ~replace(.x,
             tag %in% bad.tags,
             NA)))

```

## 4 Pivot data
Use the below code to read in the Soils data from the *carData* package
```{r}
# Load the example data 

soil <- carData::Soils    # load example data
```

* print the first few lines of data in "soil"

* Pivot the data so that columns Ca - Na are contained in one column called nutrients (again there are two possible solutions (really more than that but two I expect people to use))
```{r echo=FALSE}

#See what variables it contains...
head(soil)  

# Use 'tidyverse' to reshape the data 

soil_nut <- pivot_longer(soil, 
                         cols = c("Ca","Mg","K", "Na"), 
                         names_to = "nutrient",
                         values_to = "value" )

soil_nut

# alternatively
soil_nut <- pivot_longer(soil, 
                         cols = Ca:Na, 
                         names_to = "nutrient",
                         values_to = "value" )

soil_nut

```

## 5 Join Data

If you haven't already download the 3 bobcat data files added to the course after 12 January 2024 and save them to the data/raw folder

[Bobcat collection data for Purrr (bobcat_collection_data.csv)](data/raw/bobcat_collection_data.csv)

[Bobcat necropsy data for Purrr (bobcat_necropsy_only_data.csv)](data/raw/bobcat_necropsy_only_data.csv)

[Bobcat age data for Purrr (bobcat_age_data.csv)](data/raw/bobcat_age_data.csv)


- Read in the data files using the tidyverse **function** 
- In the same code chunk, set the column names to lowercase for all 3 data sets AND rename the 'Bobcat_ID#' column to bobcat_id   (*NOTE: this requires a lot of code repition which is annoying and does not follow best coding practice, we will learn a much better way to do this when we cover Purrr*)    
- Use the csv file names as the object names when you assign them to the environment  - Make a list with the three data sets and check their internal structure *(there are multiple ways to do this)*   
- Join the bobcat_necropsy_only_data to the bobcat_collection_data AND then in the same code chunk join the bobcat_age_data.  *You will need to use the bobcat_id column as the key when joining*   
- Print the summary of your data to check that it worked   
```{r echo=FALSE, message=FALSE}

# read in data files
bobcat_collection_data <- read_csv('data/raw/bobcat_collection_data.csv') %>% 
  
  # set names to lowercase
  set_names(
        names(.) %>% 
          tolower()) %>%
  
  # change bobcats id# to better name
      rename(.,
             'bobcat_id' = 'bobcat_id#')

bobcat_necropsy_only_data <- read_csv('data/raw/bobcat_necropsy_only_data.csv') %>% 
  
  # set names to lowercase
  set_names(
        names(.) %>% 
          tolower()) %>%
  
    # change bobcats id# to better name
      rename(.,
             'bobcat_id' = 'bobcat_id#')

bobcat_age_data <- read_csv('data/raw/bobcat_age_data.csv') %>% 
  
  # set names to lowercase
  set_names(
        names(.) %>% 
          tolower()) %>%
  
    # change bobcats id# to better name
      rename(.,
             'bobcat_id' = 'bobcat_id#')

# make a list and check internal structure 

# option 1 - nested code
str(list(bobcat_collection_data,
    bobcat_necropsy_only_data,
    bobcat_age_data))

# option 2 - with dplyr
list(bobcat_collection_data,
    bobcat_necropsy_only_data,
    bobcat_age_data) %>% 
  
  str(.)


# join data

bobcat_data_joined <- bobcat_collection_data %>% 
  
  # join necropsy data
  left_join(bobcat_necropsy_only_data,
            by = 'bobcat_id') %>% 
  
  # join age data
  left_join(bobcat_age_data,
            by = 'bobcat_id')


# print summary
summary(bobcat_data_joined)

```

