---
title: "Assignment 2"
author: "Working with data"
date: "Due 2 February 2024"
output: 
  html_document:
    code_folding: show
    theme: journal
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

#  ES482 R labs   
#      University of Victoria, Victoria BC Canada             

#  Assignment 2: working with data

library(tidyverse)
```

## Submission
Please email me a copy of your R script for this assignment by start of class on **Friday February 2nd** with your first and last name followed by assignment 2 as the file name (e.g. 'marissa_dyck_assignment2.R')

> You should always be following best coding practices (see Intro to R module 1) but especially for assingment submissions. Please make sure each problem has its own header so that I can easily navigate to your answers and that your code is well organized with spaces as described in the best coding practices section and comments as needed. 

## 1 Export data
Save the altered turtles data as a comma separated file to the *data/processed* folder in your working directory using the 'readr' package and name it 'turtles_tidy'
```{r echo=FALSE, message=FALSE}
# code for altering turtle data
turtles.df <- read_delim('data/raw/turtle_data.txt',
                         delim = '\t') %>% 
  
   # set column names to lowercase
  set_names(
    names(.) %>%  
      tolower()) %>% 
  
  # rename columns to shorter names
  rename(tag = tag_number,
         c_length = carapace_length,
         h_width = head_width) %>% 
  
  # change sex to factor
  mutate(sex = as.factor(sex))
```

```{r class.source = 'fold-hide'}

write_csv(turtles.df,
          'data/processed/turtles_tidy.csv')

```


Download the [brown bear damage data (bear_2008_2016.csv)](data/raw/bear_2008_2016.csv) and save it in the data folder. 

## 2 Import and rename data
* Import the brown bear dataset using the appropriate function from the 'readr' package and save it as bear_data.    
* In the same code chunk set the column names to lowercase and    
* change dist_to_town to m_to_town and dist_to_forest to m_to_forest (m for meters)    
* Finally, use one of the functions we've covered to view/print your data to check that it worked.

```{r class.source = 'fold-hide'}

bear_data <- read_csv('data/raw/bear_2008_2016.csv') %>% 
  
  # set column names to lowercase
  set_names(
    names(.) %>%  
      tolower()) %>% 
  
  # rename columns to include units
  rename(m_to_town = dist_to_town,
         m_to_forest = dist_to_forest)
  

head(bear_data) 
```

## 3 Change variable type
* Check the internal structure of the bear dataset    
* Change 'targetspp' from **character** to **factor** using *dplyr* make sure to overwrite the previous bear data so the changes are saved    
* Use the 'levels()' **function** to check if this worked. HINT: Use the help file if you aren't familiar with 'levels()'

```{r class.source = 'fold-hide'}

str(bear_data)

# change targetspp to factor dplyr
bear_data <- bear_data %>% 
  mutate(targetspp = as.factor(targetspp))

levels(bear_data$targetspp)
```

## 4 Subsetting
* Now that we know there are 3 livestock types (groups), subset the data to include only the rows for 'ovine' 
* In the same code chunk select the columns damage, year, month, bear_abund, landcover_code, and altitude
* Save this to your **environment** as 'bear_sheep_data'
```{r class.source = 'fold-hide'}

bear_sheep_data <- bear_data %>% 
  
  # return only rows for sheep
  filter(targetspp == 'ovine') %>% 
  
  # select specified columns
  select(damage:month, bear_abund:altitude) # most parsimonious way to do it but you could have also done


# select(damage, year, month, bear_abund, landcover_code, altitude)

# or 

# select(1:3, 7:9)

head(bear_sheep_data)
```

## 5 Summarise 
* Using the bear_sheep_data and the `summarise()` **function**, calculate the mean, sd and SE (the formula for SE is sd / sqrt(n)) for altitude. *This last part might be tricky at first but give it a try and remember you can always google things and check the 'help' files; and if that fails 'phone a friend or me'*

```{r class.source = 'fold-hide'}

bear_sheep_data %>% 
  summarise(mean_alt = mean(altitude),
            sd_alt = sd(altitude),
            se_alt = sd_alt/sqrt(length(altitude))) # you could also look up the length for altitude in the console and type the number here directly but this is more flexible and readable

# or another possible answer

bear_sheep_data %>% 
  summarise(mean_alt = mean(altitude),
            sd_alt = sd(altitude),
            se_alt = sd_alt/sqrt(n()))

```