---
title: "Assignment 1"
author: "Base R"
date: "Due 26 January 2024"
output: 
  html_document:
    code_folding: show
    theme: journal
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE)
```

```{r echo=FALSE}

#  ES482 R labs   
#      University of Victoria, Victoria BC Canada             

#  Assignment 1: base R

```

## Submission
Please email me a copy of your R script for this assignment my start of lab on Friday January 26th with your first and last name followed by assignment 1 as the file name (e.g. 'marissa_dyck_assignment1.R')

> You should always be following best coding practices (see Intro to R module 1) but especially for assingment submissions. Please make sure each problem has it's own header so that I can easily navigate to your answers and that your code is well organized with spaces as described in the best coding practices section and comments as needed. 

## 1 Create a vector
Create a vector called 'myvec' using any of the methods you learned with numbers 1 to 10. *Note there are multiple ways to do this*. 

```{r echo=FALSE}

# answer 1 (the most parsimonious)
myvec <- 1:10

# answer 2
myvec <- c(1, 10)

# answer 3 (time consuming)
myvec <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)

# answer 4 (unecessarily complicated for this exercise but works)
myvec <- seq(1, 10, by = 1)
```

## 2 Create a matrix with `rbind()`
Create a 3 row by 2 column matrix named 'mymat'. Use the `rbind()` **function** to bind the following three rows/vectors together:

```
c(1,4) 
c(2,5) 
c(3,6)
```

```{r echo=FALSE}

rbind(
c(1,4),
c(2,5),
c(3,6)
)

```

## 3 Extracting objects from data
Get the names of columns in the data frame you created earlier 'mydf'. *Hint see the R functions to explore data section*. Then extract all rows for column 5 by name, do the same thing using the element position e.g. `[]`
```{r echo=FALSE}
# code to create mydf
my.data <- data.frame(
  Obs.Id = 1:100,
  Treatment = rep(c("A","B","C","D","E"),
                  each = 20),
  Block = rep(1:20,
              times = 5),
  Germination = rpois(100,
                      lambda = rep(c(1,5,4,7,1),
                                   each = 20)),   # random poisson variable
  AvgHeight = rnorm(100,
                    mean = rep(c(10,30,31,25,35,7),
                               each = 20))
)

mydf = my.data[21:30, ]  

# answer for problem 1
names(mydf)

mydf$AvgHeight

mydf[, 5]

```

## 4 Create a matrix
Create a new matrix called 'mymat2' that includes all the data from columns 3 to 5 of data frame mydf. *HINT: use the `as.matrix()` function to coerce a data frame into a matrix. Since we didn't cover this function you may need to look it up in the help files.*

> Note your values for some columns may be slighly different since the code to create mydf uses random number generators. 
```{r echo=FALSE}

as.matrix(mydf[, 3:5])

```


## 5 Create a list
Create a list named 'mylist' that is composed of a    
  - vector: 1:3,     
  - a matrix: matrix(1:6, nrow = 3, ncol = 2),      
  - and a data frame: data.frame(x =c (1, 2, 3), y = c(TRUE, FALSE, TRUE), z = c("a", "a", "b")).
  
```{r echo=FALSE}

#create an empty list
mylist <- list()

# add a vector of 1 to 3 to the list
mylist[[1]] <- 1:3

# add a matrix to the list
mylist[[2]] <- matrix(1:6,
                      nrow = 3,
                      ncol = 2)

# add a data frame to the list
mylist[[3]] <- data.frame(x = c(1, 2,3 ),
                          y = c(TRUE, FALSE, TRUE),
                          z = c("a", "a", "b"))

# print the list
mylist

```

## 6 Extracting objects from lists
Extract the second and third observation from the 1st column of the data frame in 'mylist' (the list created above). 

```{r echo=FALSE}

# call mylist then reference the position of the data frame in the list [[3]], then the column in the data frame [[1]], and finally the observations within that columns c(2, 3)
mylist[[3]][[1]][c(2, 3)]

```
