---
title: "Assignment 3"
author: "Data manipulation"
date: "Due 16 February 2024"
output: 
  html_document:
    code_folding: show
    theme: journal
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

#  ES482 R labs   
#      University of Victoria, Victoria BC Canada             

#  Assignment 3: Data manipulation

library(tidyverse)
```

## 1 Remove NAs
Using the turtles_tidy data, remove all rows with NAs and create a new dataset from this called "turtles_no_na" 

```{r echo=FALSE, message=FALSE}

# code to read in turtles data from earlier
turtles_tidy <- read_csv('data/processed/turtles_tidy.csv') %>% 
  
  # change sex to a factor
  mutate(sex = as.factor(sex),
         sex = recode(sex, 
                       fem = 'female'))

# answer to Q1
turtles_no_na <- turtles_tidy %>% 
  
  # remove rows with NA
  na.omit()

turtles_no_na
```

## 2 Make a new variable
Using the turtles_no_na data, make a new variable called "size_class" based on the "weight" variable using case_when whereby

* weights less than 3 are juvenile

* weights greater than 6 are adult

* weights between 3 and 6 are subadult

(There are multiple ways to do this but they will yield the same answer)
```{r echo=FALSE}

turtles_tidy <- turtles_tidy %>% 
  mutate(size.class = case_when(
    weight < 3 ~ "juvenile",
    weight > 6 ~ "adult",
    TRUE  ~ "subadult"
  ))

turtles_tidy$size.class

# alternatively
turtles_tidy <- turtles_tidy %>% 
  mutate(size_class = case_when(
    weight < 3 ~ "juvenile",
    weight > 6 ~ "adult",
    weight >= 3 & weight <= 6 ~ 'subadult'
  ))

turtles_tidy$size_class
```


## 3 Replace values with NA
In the turtles_tidy data replace ALL variable values for tags 104 and 105 with NAs
```{r echo=FALSE}
# list of tags we do not trust the data for
bad.tags <- c(104,105)

turtles_tidy %>% 
  mutate(
    sex = replace(sex,
                  tag %in% bad.tags,
                  NA),
    c_length = replace(c_length,
                       tag %in% bad.tags,
                       NA),
    h_width = replace(h_width,
                      tag %in% bad.tags,
                      NA),
    weight = replace(weight,
                     tag %in% bad.tags,
                     NA)
  )

# or... use some more tidyverse helper functions and tricks!

turtles_tidy %>% 
  mutate(across(
    c("sex","c_length","h_width", "weight"),
    ~replace(.x,
             tag %in% bad.tags,
             NA)))

```

## 4 Pivot data
Use the below code to read in the Soils data in the carData package
```{r}
# Load the example data 

soil <- carData::Soils    # load example data
```

* look at the first few lines of data in "soil"

* Pivot the data so that columns Ca - Na are contained in one column called nutrients (again there are two possible solutions (really more than that but two I expect peopl to use))
```{r echo=FALSE}

#See what variables it contains...
head(soil)  

# Use 'tidyverse' to reshape the data 

soil_nut <- pivot_longer(soil, 
                         cols = c("Ca","Mg","K", "Na"), 
                         names_to = "nutrient",
                         values_to = "value" )

soil_nut

# alternatively
soil_nut <- pivot_longer(soil, 
                         cols = Ca:Na, 
                         names_to = "nutrient",
                         values_to = "value" )

soil_nut

```
