---
title: "Camera Data"
author: 'Module 1: Camera data organization and exploration'
date: "6 February 2024"
output:
  html_document:
    code_folding: show
    theme: journal
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
```

```{r echo = FALSE, message=FALSE, warning=FALSE}

#  ES482 R labs   
#      University of Victoria, Victoria BC Canada             

#  Camera Data
# Module 1: Camera data organization and exploration 


```

## Materials

### Scripts

1. Click [here](mod_1_cam_data.R) to download the script! Save the script to the 'scripts'' folder in your project directory you set up in the previous module.

2. Load your script in RStudio. To do this, open RStudio and click the **files** window and select the scripts folder and then this script.

### Cheat sheets

There are no cheat sheets specific to this module but don't forget the ones you've already printed for previous modules!

### Website

Part of the material we will cover in this module is based off standardized R scripts and procedures developed by ACME colleagues at the [WildCam Network](https://wildcams.ca/protocols/)
